---
title: "Homework 2 (in R)"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---
1. Consider the sequence $s$ defined by:

$$s_n = n^2 - 3n + 3,\ for\ n \geq 1$$
  
  a) Find 
  
  $$\sum_{i=1}^4 S_i = \sum_{i=1}^4 (i^2 - 3i + 3) = $$
```{r}
  S <- function(i) {
    i^2-(3*i)+3
  }

  sum(sapply(1:4, S))
```

  b) Find

$$\sum_{k=3}^5 S_k$$
```{r}
  sum(sapply(3:5, S))
```

Also, using the sequences $y$ and $z$ defined by:

$y_n = 2^n - 1, z_n = n(n - 1)$

  c) Find
$$\Big(\sum_{i=1}^3 y_i\Big)\Big(\sum_{i=1}^3 z_i\Big)$$
```{r}
  y <- function(n) {n^2 - 1}
  z <- function(n) {n * (n - 1)}
  
  sum(sapply(1:3, y)) * sum(sapply(1:3, z))
```


2. Consider the values:

 $x_1 = 1, x_2 = 3, x_3 = 5, x_4 = 7, x_5 = 10$
 $x_6 = 4, x_7 = 7, x_8 = 8, x_9 = -7, x_{10} = 6$

```{r}
x = c(1,3,5,7,10,4,7,8,-7,6)
```

Compute the mean and variance of the above values. Recall that
  the sample mean $(\bar{x})$ and sample variance $(s^2)$ are defined as:

  a) Sample Mean

$$\bar{x} = \frac{1}{n}\sum_{i=1}^nx_i = $$

```{r}
sample_mean <- function(x) {1/length(x) * sum(x)}
sample_mean(x)
```

  b) Sample Variance
  
$$s^2 = \frac{1}{n - 1} \sum_{i=1}^n (x_i - \bar{x})^2$$

```{r}
sample_variance <- function(x){ (1/(length(x) - 1)) * sum((x - sample_mean(x))^2) }
sample_variance(x)
```


3. Evaluate the following double summations:

  a) Find
$$\sum_{i=1}^4 \sum_{j=2}^4 (i - j)$$

```{r}
doubleSum <- function(iRange, jRange, fn) {
  double_summation = 0
  
  for(i in iRange) {
    for(j in jRange) {
      double_summation = double_summation + fn(i, j)    
    }
  }
  
  double_summation
}

doubleSum(1:4,2:4,function(i, j) {i - j})
```
  b) Find
$$\sum_{j=1}^4 \sum_{i=2}^4 (i - j)$$
```{r}
doubleSum(2:4,1:4,function(i, j) {i - j})
```


4) Consider the following points:

$$\begin{align}
 (x_1,y_1)&=(2,2),\\
 (x_2,y_2)&=(2,4),\\
 (x_3,y_3)&=(3,4),\\
 (x_4,y_4)&=(4,3),\\
 (x_5,y_5)&=(4,4),\\
 (x_6,y_6)&=(4,5),\\
 (x_7,y_7)&=(6,4),\\
 (x_8,y_8)&=(6,5)\\
\end{align}$$

Also Consider the following linear functions:

$$f_1(x) = \frac{1}{2}x+1$$
$$f_2(x) = \frac{1}{2}x+2$$


$$f_3(x) = 4$$

Evaluate which of the three linear functions is a better linear regression function for the given points. Let $f$ be any of the functions. Then for each function you will have to compute the value: 

$$SE = \sum_{i=1}^8(y_i - f(x_i))^2$$

Then you will choose the best regression function (of the three) as the one with the smallest value for $SE$.

```{r}
x <- matrix(data = c(2, 2, 3, 4, 4, 4, 6, 6, 2, 4, 4, 3, 4, 5, 4, 5), nrow = 8, ncol = 2)
```
  
$SE_{f_1(x)} =$
```{r}
sum((x[,2] - sapply(x[,1],function(n){n/2+1}))^2)
```
  
$SE_{f_2(x)} =$
```{r}
sum((x[,2] - sapply(x[,1],function(n){n/2+2}))^2)
```
  
$SE_{f_3(x)} =$
```{r}
sum((x[,2] - sapply(x[,1],function(n){4}))^2)
```

$$\therefore \text{The best linear regression function is: } f_2(x)$$